---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# AQPrediction

<!-- badges: start -->
<!-- badges: end -->

The goal of this repo is to demonstrate spatio-temporal prediction models to estimate levels of air pollution.

The input dataset is an Excel [file](https://drive.google.com/open?id=14GvlMCBLR9FQ_BsCeg0tj-pb1AadXr3G3gKPnrfm8gg) provided as part of the [OpenGeoHub Summer School 2019](https://opengeohub.org/summer_school_2019).

We'll use these packages

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(sf)
  library(ranger)
  library (CAST)
  library(caret)
})
```

And read-in the input data as follows

```{r, warning=FALSE, message=FALSE}
train = readxl::read_excel("SpatialPrediction.xlsx", sheet = 1)
covar = readxl::read_excel("SpatialPrediction.xlsx", sheet = 2)
locat = readxl::read_excel("SpatialPrediction.xlsx", sheet = 3)
# times = readxl::read_excel("SpatialPrediction.xlsx", sheet = 4) # what is this? -->Their test set. They will use to check our model
targt = readxl::read_excel("SpatialPrediction.xlsx", sheet = 5)
wind = readRDS ("https://github.com/Robinlovelace/AQPredict2/blob/master/ERA5_24h_uvWind_precip.RDS") ## CHECK if the path is correct.
```

The objective is to fill the NA values in the `targt` data:

```{r}
targt[1:3]
```

Let's do some data cleaning and plot the data:

```{r}
d = inner_join(train, covar)
d = inner_join(d, locat)
d = merge(d, wind, by.x=c("id", "time"), by.y=c("monitors.id", "date")) ### inner_join to join using two columns????
class(d$time)
d$dow  <- weekdays.POSIXt(d$time) # I included in the wind file
d$hour <- hour(d$time) # I included in the wind file
head(d); nrow(d)

dsf = sf::st_as_sf(d, coords = c("X", "Y"), crs = 4326)
summary(dsf)
mapview::mapview(dsf %>% sample_n(1000))
```

```{r}
## FORMULA ##
formula = as.formula(PM10 ~  humidity + temperature + (X*Y) + u10 + v10 + tp + dow + hour)
```
A simple model:

```{r}
##\\LINEAR//##
m = lm(formula, data = d)
## Are all the predictors significant? Let's check!
summary(m)

lm.df = d
lm.df$pred = predict(m, newdata = lm.df)
lm.random = lm(PM10 ~ pred, data=lm.df)
summary(lm.random)
plot(lm.df$PM10, lm.df$pred)
cor(lm.df$PM10, lm.df$pred)^2
```
A simple linear model can explain ~19% of the variability in PM10 levels, not great!

Lets try a fancy model: Random Forest

```{r}
##\\RANDOM FOREST//##

tunegrid = expand.grid(.mtry=c(3), .splitrule = "variance", .min.node.size = c(5))

#@ mtry = 3 (Number of variables to possibly split at in each node. Default is the (rounded down) square root of the number variables).
#@ min.node.size = 5  (Minimal node size. Default 1 for classification, 5 for regression, 3 for survival, and 10 for probability).
#@ splitrule = variance (Splitting rule. For regression "variance", "extratrees" or "maxstat" with default "variance").

## CV TYPE RANDOM
ranger.model = caret::train (formula, 
                                data= d, 
                                method="ranger", 
                                metric="RMSE",
                                num.trees = 50,
                                tuneGrid=tunegrid, 
                                trControl=trainControl(method="cv", number=10, 
                                                      allowParallel=TRUE))  
ranger.model
#  RMSE     Rsquared  MAE     
#5.989361  0.882534  3.577604

## CV TYPE SPACE
space.cv = CreateSpacetimeFolds(d, spacevar = "id")
  
ranger.space = caret::train (formula, 
                                data= d, 
                                method="ranger", 
                                metric="RMSE",
                                num.trees = 50,
                                tuneGrid=tunegrid,
                                trControl=trainControl(method="cv", index = space.cv$index,
                                                      number=10, allowParallel = TRUE))

ranger.space 
#  RMSE     Rsquared   MAE     
#11.58011  0.5467433  7.581717

space.df = d
space.df$pred = predict(ranger.space, space.df)
lm.space = lm(PM10 ~ pred, data=space.df)
summary(lm.space)


##CV TYPE TIME
time.cv = CreateSpacetimeFolds(d, spacevar = "time")
  
ranger.time = caret::train (formula, 
                                data= d, 
                                method="ranger", 
                                metric="RMSE",
                                num.trees = 50,
                                tuneGrid=tunegrid,
                                trControl=trainControl(method="cv", index = time.cv$index,
                                                      number=10, allowParallel = TRUE))
                                                      
ranger.time
#  RMSE      Rsquared   MAE     
# 6.277938  0.8732086  3.874261

time.df = d
time.df$pred = predict(ranger.time, time.df)
lm.time = lm(PM10 ~ pred, data=time.df)
summary(lm.time)

## CV TYPE SPACE-TIME 
spacetime.cv = CreateSpacetimeFolds(d, spacevar = "id",timevar ="time")
  
ranger.spacetime = caret::train (formula, 
                                data= d, 
                                method="ranger", 
                                metric="RMSE",
                                num.trees = 50,
                                tuneGrid=tunegrid,
                                trControl=trainControl(method="cv", index = spacetime.cv$index,
                                                      number=10, allowParallel = TRUE))
                                                      
ranger.spacetime
#  RMSE      Rsquared   MAE     
# 9.591133  0.6802324  5.760715

spacetime.df = d
spacetime.df$pred = predict(ranger.spacetime, spacetime.df)
lm.spacetime = lm(PM10 ~ pred, data=spacetime.df)
summary(lm.spacetime)
```

To get the result let's see what is in the target table:

```{r}
targt_joined = inner_join(targt, locat)
targt_joined = inner_join(targt_joined, covar)
targt_joined = merge(targt_joined, wind, by.x=c("id", "time"), by.y=c("monitors.id", "date")) ## inner_join to join using two columns????

summary(targt_joined)
summary(targt$id %in% d$id)
summary(targt$id %in% covar$id)
summary(targt$id %in% locat$id)
summary(targt$id %in% wind$monitors.id)
```
Now we can predict the missing values:

```{r}
p = predict(ranger.spacetime, newdata = targt_joined)
summary(p)
write.csv(data.frame(p = p), "p.csv")
```

```{r, echo=FALSE}
# system.time(x = 1:1e9)
# system.time({x = 1:1e9})
```

